---
title:  |  
        | FIE401 - First assignment 
        | M&A and Method of Payment
author: 'Group 8'
date: '`r Sys.Date()`'
output: 
  pdf_document:
    fig_caption: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Abstract

In this report we intend to answer the research question: “Do mergers destroy value for bidder shareholders depending on the method of payment?” We have used the provided file “CAR_M&A.Rdata” which contains 5154 deals over a period from 1990 to 2014 to investigate the impact of payment method. Our findings on this topic are that when a bidding company fully uses stocks as currency for acquisitions it receives a negative announcement effect on the wealth if the target company is a public company. In contrast, the effect tends to be the opposite when the target company is private.  

```{r, results='asis', message = FALSE}
# SETUP ------------------------------------------------------------------------
setwd('C:/Users/steef/Documents/NHH/FIE401 Financial Econometrics/Fin-Econ-Assignment-1')
load('CAR_M&A.RData')

# Libraries
library(ggplot2)
library(RColorBrewer)
library(tidyverse)
library(stargazer)
library(corrplot)
library(gridExtra) # for combining graphs
library(lubridate) # for ymd function 

# Functions, colors and themes
mytheme <- 
  theme_minimal() +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank())
```

# Exploratory analysis and outlier detection

```{r summary-stats, echo=FALSE, results='asis'}
# STEP 1 -----------------------------------------------------------------------
#       (Summary statistics & Exploratory analysis)

# Variables public and private contain the same information
# change to date variable
CAR_MA$yyyymmdd <- ymd(CAR_MA$yyyymmdd)

# Binary variables are stored as numerical
binary.vars <- c('public', 'private', 'tender_offer', 'all_stock', 'hostile',
                 'horz')

stargazer(CAR_MA[!names(CAR_MA) %in% c('yyyymmdd','yyyy',binary.vars)], 
          type = 'latex', header = FALSE, 
          summary.stat = c('min','p25','mean','p75', 'max','sd'))
```

**Comments on the Table 1**
Even a simple summary of the data shows that there are outliers in the data, for example by large differences in mean and median or between maximum value and 3rd quantile. Outliers are a significant problem in empirical research because they can cause biased regression coefficients. Two obvious outliers from the summary are deal value and bidder size.

There are statistical tools to find outliers, one of the most famous methods in regression setting is Cook`s distance, which basically is a summary on how much the regression outcome changes when specific observations are removed.  We however will rely on plots, which is also a good way to find outliers in the data. In this case we plotted a scatterplot of the variable to see whether observations are distant from others. This can be clearly seen with the deal value and bidder size.  Overall, there seem to be other “smaller” outliers. 

```{r}
# Are there outliers? Evaluate if you should winsorize the data:
# Deal_value:
plot1 <- 
  CAR_MA %>% 
  ggplot(aes(x = yyyy,y = deal_value))+
  geom_point(color = brewer.pal(11,'PiYG')[10]) +
  geom_point(data = CAR_MA[2558,], color = brewer.pal(11,'PiYG')[2]) +
  mytheme +
  labs(x = 'Year',
       y = 'Deal value')

plot2 <- 
  CAR_MA %>%   
  ggplot(aes(deal_value))+
  geom_histogram(bins = 30,
                 fill = brewer.pal(11,'PiYG')[2]) +
  mytheme +
  labs(x = 'Deal value',
       y = 'Frequency')
```

```{r fig.width = 9, fig.height = 5,fig.cap="\\label{fig:figs}Scatterplot and histogram of deal value"}
grid.arrange(plot1, plot2, ncol=2)
```

**Comments to figure 1**


```{r}
# bidder_size
plot3 <- 
  CAR_MA %>% 
  ggplot(aes(x = yyyy, y = bidder_size))+
  geom_point(color = brewer.pal(11,'PiYG')[10]) +
  mytheme +
  labs(x = 'Year',
       y = 'Bidder size')

plot4 <- 
  CAR_MA %>%   
  ggplot(aes(x = factor(0), bidder_size))+
  geom_boxplot(alpha = 0.5, color = brewer.pal(11,'PiYG')[2]) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  mytheme +
  labs(x = '',
       y = 'Bidder size')
```

```{r fig.width = 9, fig.height = 5,fig.cap="\\label{fig:figs}Scatterplot and boxplot of bidder size"}
grid.arrange(plot3,plot4,ncol=2)
```

**Comments to figure 2**


```{r}
plot5 <- 
  CAR_MA %>% 
  ggplot(aes(x=yyyy, y = bidder_mtb))+
  geom_point(color = brewer.pal(11,'PiYG')[10]) +
  mytheme +
  labs(x = 'Year',
       y = 'Bidder market-to-book ratio')

# bidder book to market 
plot6 <- 
  CAR_MA %>%   
  ggplot(aes(x=factor(0),bidder_mtb))+
  geom_boxplot(alpha = 0.5, color = brewer.pal(11,'PiYG')[2]) +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
  mytheme +
  labs(x = '',
       y = 'Bidder market-to-book ratio')
```

```{r fig.width = 9, fig.height = 5,fig.cap="\\label{fig:figs}Scatterplot and boxplot of bidder book-to-market ratio"}
grid.arrange(plot5,plot6,ncol=2)
```

**Comments to figure 3**
Seems suspicious to be an error but we believe the reason for high values only in year 2000 is the dotcom bubble 


```{r fig.width = 9, fig.height = 5,fig.cap="\\label{fig:figs}Deal size histogram before and after log-transformation"}
# Deal with outliers
CAR_MA <- 
  CAR_MA %>%
  rename(year = yyyy) %>% 
  slice(-(2558)) %>%  # to get rid of extreme outlier
  mutate(deal_value = log(deal_value),
         bidder_size = log(bidder_size))

plot2.0 <- 
  CAR_MA %>%   
  ggplot(aes(deal_value))+
  geom_histogram(bins = 30,
                 fill = brewer.pal(11,'PiYG')[10]) +
  mytheme +
  labs(x = 'Deal value',
       y = 'Frequency')

grid.arrange(plot2, plot2.0, ncol=2)
```


# Descriptive table 1 
```{r results='asis'}
# STEP 2 -----------------------------------------------------------------------
#       (Average values by year)

# descriptive table 1:
summary1 <- 
  CAR_MA %>% 
  arrange(yyyymmdd) %>% 
  group_by(year) %>%
  mutate(share_private = mean(private),
         share_stock = mean(all_stock)) %>% 
  summarise(avg_deal_size = round(mean(deal_value), digits= 3),
            avg_bidCAR = round(mean(carbidder), digits=3),
            avg_share_private = round(mean(share_private), digits=3),
            avg_share_stock = round(mean(share_stock), digits = 3))

stargazer(summary1, summary = FALSE, type = 'latex', header = FALSE)
```
**Comments descriptive table 1**

This table displays the average deal size, carbidder, share of deals with private firms, and share of deals fully paid in stock per year in the data. Key insights in this table: 
- The average deal size went up in the period. Especially in 2014. This can indicate some extreme values that should be winsorized out. But is also to some extent expected because of economic growth. 
- The abnormal returns have varied from –0,0212 to 0,0425. Returns have a straightforward effect on wealth, therefore give good information about research questions about the wealth of bidder shareholders 
- The share of deals with private targets have had a relatively stable percentage over the period under investigation.
- The biggest change in the table is the share of deals fully paid in stock. This share was stable from 1990-1999 (around 50%), but from 1999-2014 we see a steep decrease in this share. The period ended in just 17%.  This is beneficial to find an insight of the methods of payment and answer that research question. 

# Descriptive table 2 
```{r}
# STEP 3 -----------------------------------------------------------------------
#       (Average values by method of payment & t-tests)

vars.for.model <- names(CAR_MA[!names(CAR_MA) %in% c('yyyymmdd','yyyy', 
                                                     'private', 'all_stock')])

difference.means <- data.frame(matrix(NA,    # Create empty data frame
                                      nrow = 4,
                                      ncol = length(vars.for.model)))

row.names(difference.means) <- c('all.stock', 'not.all.stock', 'differece', 
                                 't-test')
names(difference.means) <- vars.for.model

difference.means[1,] <- lapply(CAR_MA[CAR_MA$all_stock == 1,
                                      names(CAR_MA) %in% vars.for.model], mean)

difference.means[2,] <- lapply(CAR_MA[CAR_MA$all_stock == 0,
                                      names(CAR_MA) %in% vars.for.model], mean)

difference.means[3,] <- difference.means[1,] - difference.means[2,]

for (v in vars.for.model){
  pval <- t.test(CAR_MA[CAR_MA$all_stock == 1,v],
                 CAR_MA[CAR_MA$all_stock == 0,v])$p.value
  difference.means[4,v] <- pval
}
```
** Comments descriptive table 2 **
- test
- test
- test
- 
# Regression

## Correlation Plot
```{r fig.width = 9, fig.height = 5,fig.cap="\\label{fig:figs}Correlation plot"}



# STEP 4 -----------------------------------------------------------------------
#       (Regression table)

# Check correlations for potential multicolliniarity
correlations <- cor(CAR_MA[!names(CAR_MA) %in% 'yyyymmdd'])
corrplot(correlations, col = COL2('PiYG'), tl.col = 'black')
```
** Comments on Correlation Plot**
As we can observe, the correlation between bidder size and deal value is high relative to many of the other observations. As bigger firms tend to have higher incomes and sometimes larger cash reserves, they can also buy other big firms, resulting in higher values on deals. 

The variables “public” and “private” have a highly negative correlation mainly due to the fact that they represent the direct opposite of each other. 

## Regression table
```{r results='asis'}
# run regressions 

# Only public - no controls
model1 <- 
  lm(carbidder ~ all_stock, CAR_MA[CAR_MA$public == 0,]) 
# summary(model1)

# Only private - no controls
model2 <- 
  lm(carbidder ~ all_stock,
     CAR_MA[CAR_MA$public == 1,]) 
# summary(model2)


# Full sample - no controls
model3 <- 
  lm(carbidder ~ all_stock +
                 public +
                 I(all_stock*public),
                 CAR_MA) 
# summary(model3)


# Only public - controls
model4 <- 
  lm(carbidder ~ all_stock +
       deal_value +
       bidder_size +
       bidder_mtb + 
       run_up_bidder +
       bidder_fcf +
       bidder_lev +
       sigma_bidder + 
       relsize +
       horz +
       tender_offer +
       year + # ********************************************************** Add as FE
       hostile,
     CAR_MA[CAR_MA$public == 0,])
# summary(model4)


# Only private - controls
model5 <- 
  lm(carbidder ~ all_stock +
       deal_value +
       bidder_size +
       bidder_mtb + 
       run_up_bidder +
       bidder_fcf +
       bidder_lev +
       sigma_bidder + 
       relsize +
       horz +
       tender_offer +
       year + # ********************************************************** Add as FE
       hostile,
     CAR_MA[CAR_MA$public == 1,])
# summary(model5)


# Full sample - controls
model6 <-
  lm(carbidder ~ all_stock +
                 public +
                 I(all_stock*public) +
                 deal_value +
                 bidder_size +
                 bidder_mtb +
                 run_up_bidder + 
                 bidder_fcf +
                 bidder_lev +
                 sigma_bidder +
                 relsize +
                 horz + 
                 tender_offer +
                 year + # ********************************************************** Add as FE
                 hostile,
                 CAR_MA)
# summary(model6)


stargazer(model1, model2, model3, model4, model5, model6,
          keep.stat = c('n','adj.rsq'),
          type = 'latex', header = FALSE, digits = 3,
          notes.append = TRUE, notes.align = 'l', font.size = 'small',
          notes = 'Notes with explanations?') # add notes ***************************************************

```
** Comments on regression table**
- Variable all_stocks always statistically significant on 1% level.
- Also economically significant, if payment procedure is done by buying all stocks ( all_stocks=1) then abnormal return of bidder increases by 0,014 (1,4%), which makes a difference with big numbers in play.  
- Looking at regressions 5 and 6 we see that different effect of all_stocks, this shows that it makes a difference whether we look at public companies or private companies.
- Therefore regression 6 is our best model in explaining the effect of all_stocks  
- However, we are not too confident that this result is unbiased, probably omitted variable bias is still a problem. We would industry fixed effects. 
- And since the adjusted Rsquare is very low (0,061) our model does not seem to fit well
- We have adjusted R^2 ranging from 0,003-0,085. A low adjusted R^2 indicated that our models does not explain much of the variation in carbidder return. If we had a relatively high R^2 (70%<), we would be able to construct an almost perfect event in which we maximized carbidder abnormal return. Since there are several factors included her, and a few other variables that could cast a light on the research question, a low adjusted R^2 is reasonable, and thus indicates that our models are consistent with reality and literature
- We think we found a relevant effect of interest but would still be completely confident about it. For example, we would have liked to include an industry variable. 
- From regression 3 we discover that if the deal is done in all stocks, the carbidder abnormal return increases by 1,4%. If the deal is done in all stock, and with a public company, this increase is neutralized by a decrease on 1,9% from the public company. This is supported by the signal effect, where if a company is set to by a large percentage, or a company in its entirety, the Share price will go up as a result of a signal that the company is doing well, and a lack of shares available in the market. This will then again result in higher total price for the “carbidder”. This supports the research question that the payment method has an effect on carbidder value. 

# Additional Questions 

## Questions 1 
** Can you include both a dummy for private and public targets in the regression?**
We cannot include both dummy variables in the regression, as we will encounter the “dummy variable trap”. The dummy variable trap is when two or more variables created by on-hot encoding are highly correlated. This causes perfect multicollinearity, which means that we can’t measure the betas correctly. Multicollinearity appears when one or more independent variables can be written as a perfect linear combination of one or more of the independent variables. The solution is by omitting one of the dummy variables. As firms either are publicly traded or private, including both dummy variables will cause multicollinearity and false betas

## Question 2 
** If you limit the sample to private deals, what’s the coefficient on the hostile variable?** 
If we limit the sample to private deals, the coefficient on the hostile variable will be zero. A hostile takeover can only happen to a public company. In a hostile takeover, the acquirer attempts to take over a company (the target company) against the wishes from the management in the target company. This can be done by going directly to the target company’s shareholders or fight to replace the management.  
