---
title: 'FIE401 - First assignment M&A and Method of Payment'
author: 'Group 8'
date: '`r Sys.Date()`'
output: 
  pdf_document:
    fig_caption: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Abstract

In this report we intend to answer the research question: “Do mergers destroy value for bidder shareholders depending on the method of payment?” We have used the provided file “CAR_M&A.Rdata” which contains 5154 deals over a period from 1990 to 2014 to investigate the impact of payment method. Our findings on this topic are that when a bidding company fully uses stocks as currency for acquisitions it receives a negative announcement effect on the wealth if the target company is a public company. In contrast, the effect tends to be the opposite when the target company is private.  

```{r, results='asis', message = FALSE}
# SETUP ------------------------------------------------------------------------
setwd('C:/Users/steef/Documents/NHH/FIE401 Financial Econometrics/Fin-Econ-Assignment-1')
load('CAR_M&A.RData')

# Libraries
library(ggplot2)
library(RColorBrewer)
library(tidyverse)
library(stargazer)
library(corrplot)
library(gridExtra) # for combining graphs
library(lubridate) # for ymd function 

# Functions, colors and themes
mytheme <- 
  theme_minimal() +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank())
```

# Question 1

```{r summary-stats, echo=FALSE, results='asis'}
# STEP 1 -----------------------------------------------------------------------
#       (Summary statistics & Exploratory analysis)

# Variables public and private contain the same information
# change to date variable
CAR_MA$yyyymmdd <- ymd(CAR_MA$yyyymmdd)

# Binary variables are stored as numerical
binary.vars <- c('public', 'private', 'tender_offer', 'all_stock', 'hostile',
                 'horz')

stargazer(CAR_MA[!names(CAR_MA) %in% c('yyyymmdd','yyyy',binary.vars)], 
          type = 'latex', header = FALSE, 
          summary.stat = c('min','p25','mean','p75', 'max','sd'))
```

**Comments on the Table 1**

```{r}
# Are there outliers? Evaluate if you should winsorize the data:
# Deal_value:
plot1 <- 
  CAR_MA %>% 
  ggplot(aes(x=yyyy,y=deal_value))+
  geom_point(color = brewer.pal(11,'PiYG')[10]) +
  geom_point(data = CAR_MA[2558,], color = brewer.pal(11,'PiYG')[2]) +
  mytheme

plot2 <- 
  CAR_MA %>%   
  ggplot(aes(deal_value))+
  geom_histogram(bins = 30,
                 fill = brewer.pal(11,'PiYG')[10]) +
  mytheme 
```

```{r fig.width = 9, fig.height = 5,fig.cap="\\label{fig:figs}Scatterplot and histogram of deal value"}
grid.arrange(plot1, plot2, ncol=2)
```

**Comments to figure 1**
just by looking at the scatterplot we indicate a huge outlier 
This could be an error but nonetheless should be removed from the dataset   
overall there seem to be other "smaller" outliers and we therefore have to winsorize or use log-transformation 
because of the shape of the historgram we decided to use log-transformations 

```{r fig.width = 9, fig.height = 5,fig.cap="\\label{fig:figs}Scatterplot and boxplot of bidder size"}
# bidder_size
plot3 <- 
  CAR_MA %>% 
  ggplot(aes(x=yyyy, y=bidder_size))+
  geom_point() +
  mytheme

plot4 <- 
  CAR_MA %>%   
  ggplot(aes(x=factor(0),bidder_size))+
  geom_boxplot()+
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  mytheme

grid.arrange(plot3,plot4,ncol=2)
```

```{r}
plot5 <- 
  CAR_MA %>% 
  ggplot(aes(x=yyyy, y=bidder_mtb))+
  geom_point() +
  mytheme

# bidder book to market 
plot6 <- 
  CAR_MA %>%   
  ggplot(aes(x=factor(0),bidder_mtb))+
  geom_boxplot()+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
  mytheme

grid.arrange(plot3,plot4,ncol=2)
# seems suspicious to be an error 
# but we believe the reason for high values only in year 2000 
# is the dotcom bubble 


# adjust data 
# adjust data 
CAR_MA <- 
  CAR_MA %>%
  rename(year = yyyy) %>% 
  slice(-(2558)) %>%                    # to get rid of extreme outlier
  mutate(deal_value = log(deal_value),
         bidder_size = log(bidder_size))

# as we see now, log-tranformation solved the outlier problem 
plot1 <- 
  CAR_MA %>% 
  ggplot(aes(x=year,y=deal_value))+
  geom_point() +
  mytheme

plot2 <- 
  CAR_MA %>%   
  ggplot(aes(deal_value))+
  geom_histogram(bins = 30,
                 fill = brewer.pal(11,'PiYG')[2]) +
  mytheme 

grid.arrange(plot1, plot2, ncol=2)



# STEP 2 -----------------------------------------------------------------------
#       (Average values by year)

# descriptive table 1:
summary1 <- 
  CAR_MA %>% 
  arrange(yyyymmdd) %>% 
  group_by(year) %>%
  mutate(share_private = mean(private),
         share_stock = mean(all_stock)) %>% 
  summarise(avg_deal_size = round(mean(deal_value), digits= 3),
            avg_bidCAR = round(mean(carbidder), digits=3),
            avg_share_private = round(mean(share_private), digits=3),
            avg_share_stock = round(mean(share_stock), digits = 3))

stargazer(summary1,summary = FALSE, type = 'latex', header = FALSE)



# STEP 3 -----------------------------------------------------------------------
#       (Average values by method of payment & t-tests)

vars.for.model <- names(CAR_MA[!names(CAR_MA) %in% c('yyyymmdd','yyyy', 
                                                     'private', 'all_stock')])

difference.means <- data.frame(matrix(NA,    # Create empty data frame
                                      nrow = 4,
                                      ncol = length(vars.for.model)))

row.names(difference.means) <- c('all.stock', 'not.all.stock', 'differece', 
                                 't-test')
names(difference.means) <- vars.for.model

difference.means[1,] <- lapply(CAR_MA[CAR_MA$all_stock == 1,
                                      names(CAR_MA) %in% vars.for.model], mean)

difference.means[2,] <- lapply(CAR_MA[CAR_MA$all_stock == 0,
                                      names(CAR_MA) %in% vars.for.model], mean)

difference.means[3,] <- difference.means[1,] - difference.means[2,]

for (v in vars.for.model){
  pval <- t.test(CAR_MA[CAR_MA$all_stock == 1,v],
                 CAR_MA[CAR_MA$all_stock == 0,v])$p.value
  difference.means[4,v] <- pval
}

# STEP 4 -----------------------------------------------------------------------
#       (Regression table)

# Check correlations for potential multicolliniarity
correlations <- cor(CAR_MA[!names(CAR_MA) %in% 'yyyymmdd'])
corrplot(correlations, col = COL2('PiYG'), tl.col = 'black')

# run regressions 

# Only public - no controls
model1 <- 
  lm(carbidder ~ all_stock, CAR_MA[CAR_MA$public == 0,]) 
summary(model1)

# Only private - no controls
model2 <- 
  lm(carbidder ~ all_stock,
     CAR_MA[CAR_MA$public == 1,]) 
summary(model2)



# Full sample - no controls
model3 <- 
  lm(carbidder ~ all_stock +
                 public +
                 I(all_stock*public),
                 CAR_MA) 
summary(model3)


# Only public - controls
model4 <- 
  lm(carbidder ~ all_stock +
       deal_value +
       bidder_size +
       bidder_mtb + 
       run_up_bidder +
       bidder_fcf +
       bidder_lev +
       sigma_bidder + 
       relsize +
       horz +
       tender_offer +
       year+
       hostile,
     CAR_MA[CAR_MA$public == 0,])
summary(model4)


# Only private - controls
model5 <- 
  lm(carbidder ~ all_stock +
       deal_value +
       bidder_size +
       bidder_mtb + 
       run_up_bidder +
       bidder_fcf +
       bidder_lev +
       sigma_bidder + 
       relsize +
       horz +
       tender_offer +
       year+
       hostile,
     CAR_MA[CAR_MA$public == 1,])
summary(model5)


# Full sample - controls
model6 <-
  lm(carbidder ~ all_stock +
                 public +
                 I(all_stock*public) +
                 deal_value +
                 bidder_size +
                 bidder_mtb +
                 run_up_bidder + 
                 bidder_fcf +
                 bidder_lev +
                 sigma_bidder +
                 relsize +
                 horz + 
                 tender_offer +
                 year+
                 hostile,
                 CAR_MA)
summary(model6)




stargazer(model1, model2, model3, model4, model5, model6,
          keep.stat = c('n','adj.rsq'),
          type = 'latex', header = FALSE, digits = 3,
          notes.append = TRUE, notes.align = 'l', font.size = 'small',
          notes = 'Notes with explanations?') # add notes ***************************************************

```

